---
title: "Fun with APIs"
author: "D Pei Wu"
date: "April 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(jsonlite)
library(dplyr)


```
```{r get trimet data}
appID <- "appID=[number]"
baseurl <- "https://developer.trimet.org/ws/v2/arrivals"
locIDs <- "locIDs=5887,5889,5890,5892"

call_1 <- paste(baseurl, "?", # "?" represents query
                locIDs, "&",
                appID, sep = "")
call_1


```

## Note on API data downloads
look for formats, IDs on the site (in this case TriMet), baseurl, format of how to make the call

generally there are wrappers already written for getting API data for major data sources but it may not have all the specific call or specific data that you need, so you might have to fall back on the long way w JSON etc

for instance there is a GTFS R package (live transit feeds) - one maintained by ropensci
there is also a package called _tidytransit_



```{r make call}
# make a GET request

get_arrivals <- httr::GET(call_1)
http_status(get_arrivals) #checks the status of the GET. if not 200 look it up


# if you ever get a data file that is lists of lists, you can use a function rlist
names(get_arrivals) 
headers(get_arrivals)

#when using lists of lists, use double brackets
parse_arrivals <- fromJSON(content(get_arrivals,  "text"))
results <- parse_arrivals[["resultSet"]]
arrivals <- results[["arrival"]]
head(arrivals)

```

## Note on Leaflet

has more options for production value: 

e.g. coding maps to make them prettier, icons, basemap layers, etc.

caveat - need to be familiar with html and/or CSS

if start using Shiny, javascript helps to understand concepts

good tutorials: jenny bryan, joe chang, others ....look them up!

also for interactive maps

## d3
wrapper for using pie charts

## Plotly
a wrapper for ggplot and plotly to make your ggplots interactive

## Regular expressions
sub:
gsub: replaces

grep: looking for a specific patter of first occurrence
grepl: replaces all occurances of patter (logical)


```{r intro to regular expressions}
# this is in base r, there is also a package called stringr 
# which may be more robust when manipulating a data frame
# icon is better than R for text mining things

biketown_raw <- biketown_2018

biketown_raw %>%
  filter(grepl("community", ignore.case = T, StartHub))

biketown_stations <- biketown_raw %>%
  mutate(station.category = if_else(grepl("community", StartHub, ignore.case = T),
                                    "Community Corral",
                                    if_else(grepl(" ", StartHub, ignore.case = T),
                                            "BIKETOWN Station", "Out of Station")))


table(biketown_stations$station.category)
```
