---
title: "Working with dates and regex"
author: "D Pei Wu"
date: "April 9, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(dplyr)
library(ggplot2)
```

## Purpose
Explore biketown data looking at the mean miles traveled per week during the summer of 2018 (06/2018 to 09/2018).

``` {r pull_data}
source("C:/Users/dpwpd/Documents/DataScienceClass/ds19-class/code/fetch_biketown.r")
get_data(start = "06/2018", end = "09/2018")
setwd("/Users/dpwpd/Documents/DataScienceClass/ds19-class/data/biketown/")

# folder <- "/Users/dpwpd/Documents/DataScienceClass/ds19-class/data/biketown/"
# filenames <- list.files(path = "C:/Users/dpwpd/Documents/DataScienceClass/ds19-class/data/biketown/", pattern = "*.csv", all.files = FALSE,
#                         full.names = FALSE,
#                         recursive = FALSE, ignore.case = FALSE)
# 
# read_csv_filename <- function(filenames) {
#   ret <- read.csv(filenames, stringsAsFactors = F,
#                   strip.white = T, na.strings = "")
#   ret$Source <- filenames
#   ret
# }
# 
# bike_raw <- plyr::ldply(filenames, read_csv_filename)

file1 <- read.csv("/Users/dpwpd/Documents/DataScienceClass/ds19-class/data/biketown/2018_06.csv", 
                  stringsAsFactors = F)
file2 <- read.csv("/Users/dpwpd/Documents/DataScienceClass/ds19-class/data/biketown/2018_07.csv", 
                  stringsAsFactors = F)
file3 <- read.csv("/Users/dpwpd/Documents/DataScienceClass/ds19-class/data/biketown/2018_08.csv", 
                  stringsAsFactors = F)
file4 <- read.csv("/Users/dpwpd/Documents/DataScienceClass/ds19-class/data/biketown/2018_09.csv", 
                  stringsAsFactors = F)

# After calling in all files use bind_rows()
bike_raw <- bind_rows(file1, file2, file3, file4)
```

```{r data_wrangling}
str(bike_raw)

# pulls date and time into one field for start and end

bike_dfl <- bike_raw %>%
  mutate(start.datetime = paste(StartDate, StartTime, sep = " "),
          end.datetime = paste(EndDate, EndTime, sep = " "))

str(bike_dfl)

# convert start.datetime and end.datetime into date-time format
# make sure to include your timezone

# check acceptable timesone values
OlsonNames()


bike_dfl$start.datetime <- mdy_hm(bike_dfl$start.datetime)
bike_dfl$end.datetime <- mdy_hm(bike_dfl$end.datetime)

# convert dDuration into a usable format
bike_dfl$Duration <- hms(bike_dfl$Duration)

# Error thrown after convertine to duration indicating NAs, 
# check raw dataframe for NAs in Duration column
sum(is.na(bike_raw$Duration))


```
'lubridate' has 3 synonymous functions but they define very different actions:
1. duration(): time spans are in unites of  seconds and there is no start date involved
2. interval(): also measured in units of seconds and provides the difference between start and end
3. period(): measures time span in units larger than seconds, is handy when accounting for DST and leap years.

```{r durationn_cal}
# calc interval
bike_dfl$interval <- interval(bike_dfl$start.datetime, bike_dfl$end.datetime)
head(bike_dfl$interval)

# calc duration
bike_dfl$duration.all <- as.duration(bike_dfl$interval)
head(bike_dfl$duration.all)

# calc period
bike_dfl$period <- as.period(bike_dfl$duration.all)
head(bike_dfl$period)

# floor.date() is helpful for aggregating
# also ceiling.date or round.date
bike_wkagg <- bike_dfl %>%
  mutate(week.datetime = floor_date(start.datetime, unit = "weeks"))
head(bike_wkagg$week.datetime)

# calculate weekly mean distance traveled


# insert bar chart w avg dist traveled by payment type
mean_dt <- bike_wkagg %>%
  group_by(week.datetime) %>%
  summarize(mean.miles = mean(Distance_Miles))


# plot results
wk_meandt_fig <- mean_dt %>%
  ggplot(aes(x = week.datetime, y = mean.miles)) +
  geom_bar(stat = "identity", fill = "orange")
wk_meandt_fig

```