---
title: "Split_apply_combine"
author: "D Pei Wu"
date: "April 9, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(modelr)
library(purrr)
library(broom)

```

## Purpose
Explore and model bike counts on 3 PDX bridges

## Our questions:
1. How much of bike use can we pick up based on seasonality? 
2. Are commuters over some bridges more weather-resistant than others?

```{r init}
getwd()
"C:/Users/dpwpd/Documents/DataScienceClass/ds19-class"
"data/"

```

```{r load data}
source("code/load_counts.r")

```

```{r join data}
bikecount_wx <- bikecounts %>%
  left_join(weather, by = c("date" = "DATE")) %>%
  select(date, bridge, total, TMIN, TMAX, PRCP)

```
dplyr joins are great! Left join above returns each row in bikecounts (left) with each amthching weather row (right), binding the weather columns by date. For more, check out tutorial on mikoontz.github

```{r explore summarizing}
bikecount_wx %>% 
  ggplot(aes(x=date, y=total,
       group = bridge, color = bridge)) +
  geom_line() + 
  # separates each bridge graph vertically
  facet_wrap(bridge ~ .)

bikecount_wx %>% 
  ggplot(aes(x=date, y=total,
       group = bridge, color = bridge)) +
  geom_line() + 
  # separates each bridge graph horizontally
  facet_grid(bridge ~ .)

bikecount_wx %>% 
  ggplot(aes(x=date, y=total,
       group = bridge, color = bridge)) +
  geom_line() + 
  # separates each bridge graph horizontally, adds yaxis limit
  facet_grid(bridge ~ .) + ylim(c(0, 10000))

```

```{r explore some summary stats}
# avg daily
bikecount_wx %>%
  group_by(bridge) %>%
  summarize(avg_daily_counts = mean(total)) %>%
  View()

# avg monthly by year
bikecount_wx %>%
  group_by(bridge, year(date), month(date)) %>%
  summarize(avg_monthly_counts = mean(total)) %>%
  View()

# daily avg monthly all years
avg_monthly <- bikecount_wx %>%
  group_by(bridge, month = month(date, label = T)) %>%
  summarize(avg_monthly_counts = mean(total))

ggplot(avg_monthly, aes(x = month, 
                        y = avg_monthly_counts,
                        color = bridge)) +
  geom_point(size = 2)


str(avg_monthly)

```

## Fitting models in R
```{r model fitting}
# models return list objects describing fit
fit <- lm(total ~ TMIN + TMAX + PRCP + bridge,
          data = bikecount_wx)
str(fit)
summary(fit)  
plot(fit)

# attempt to fit to log dist
fit2 <- lm(log(total) ~ TMIN + TMAX + PRCP + bridge,
          data = bikecount_wx)
str(fit2)
summary(fit2)  
plot(fit2)

# attempt to explore if temp stops coorelation after a max temp
fit_poly <- lm(total ~ TMIN + TMAX + I(TMAX^2) + PRCP + bridge,
   data = bikecount_wx)
summary(fit_poly)
coef(fit_poly) # could access coefs here by name or position

f <- function(x) {197 * x + -0.99* x^2} # from regression result

# stat functions extend a function over a range
p <- ggplot(data = data.frame(x=0),
            mapping = aes(x = x)) 

p + stat_function(fun=f, color="lightblue", size=2) +
  xlim(0,110) + xlab("deg F") + ylab("count")



```

```{r modelr sneak peek}
bikecount_wx <- bikecount_wx %>%
  add_predictions(fit_poly) %>%
  add_residuals(fit_poly) %>%
  mutate(pred = pred)
head(bikecount_wx)

ggplot() +
  geom_line(data = bikecount_wx, aes(x=date, y = total),
            color = "salmon") +
  geom_line(data = bikecount_wx, aes(x=date, y=pred),
           color = "blue") +
  facet_grid(bridge ~ ., scales = "free") # free removes constraint
 
```
```{r split-apply-combine}
# this nests data by bridge
bw_nested <- bikecount_wx %>%
  group_by(bridge) %>%
  nest()
str(bw_nested)
View(bw_nested)

# list factors in bw_nested 
bw_nested[[1, "bridge"]]

# fit to single bridge
lm(total ~ TMIN + TMAX + PRCP,
          data = bw_nested[[1, "data"]]) %>%
  summary()

# fit to all three bridges
fit_func <- function(df) {
  lm(total ~ TMIN + TMAX + PRCP,
     data = df)
}

fit_func2 <- function(df) {
  lm(log(total) ~ TMIN + TMAX + PRCP,
     data = df)
}

# one bridge at a time via function
fit_func2(bw_nested[[1, "data"]])

# two at a time
fits <- purrr::map(bw_nested$data[1:2], 
                   fit_func2)
fits
str(fits)

# three at a time
bw_nested <- bw_nested %>%
  mutate(fit = purrr::map(data, fit_func2),
         tidy = purrr::map(fit, tidy),
         glance = purrr::map(fit, glance))




# unpack or combine results
broom::tidy(bw_nested$fit[[1]])


# unpack for all three
bw_nested %>%
  unnest(glance) %>%
  arrange(desc(r.squared))

bw_nested %>%
  select(bridge, tidy) %>%
  unnest(tidy) %>%
  arrange(term, estimate, bridge) %>%
  View()

  
```