---
title: "Spatial Data"
author: "D Pei Wu"
date: "April 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("sf")
library(dplyr)
library(sf)
install.packages("lwgeom")
library(lwgeom)
install.packages("tidycensus")
library(tidycensus)
library(tidyr)
install.packages("ggsn")
library(ggsn)
library(mapview)
```


## Spatial manipulation with sf
API Keys:
DON'T COMMIT THIS
DON'T PUSH THIS
you can store API keys in separate files that are not linked to github or ny r projects, etc (paste in; remove it before pushing etc)
there are some packages out there that deal with API keys and encryption
there are always bots trolling on GitHub for info like this
If you do commit you basically should delete your whole repository

ggplot does mapping. let's do this!

PostGIS (Python; does GIS & mapping)

```{r tidycensus}
# this loads API key for the session * DELETE ME WHEN DONE *
# census_api_key("insert_key_here")

# object v17, loaded from CEnsus site
v17 <- tidycensus::load_variables(2017, "acs5", cache = T)
pdx_tracts_raw <- get_acs(geography = "tract", 
                          year = 2017,
                          variables = c("B19013_001",
                                        "B03002_003",
                                        "B01003_001"),
                          state = "OR",
                          county = c("Multnomah County"),
                          geometry = T) %>% # variables medinc, white NH, pop
  select(GEOID, variable, estimate, geometry)
head(pdx_tracts_raw)

```

``` {r tidycensus tidy}
pdx_tracts <- pdx_tracts_raw %>%
  spread(key = variable, value = estimate) %>%
  rename(medinc = B19013_001, white = B03002_003, 
         pop = B01003_001) %>%
  mutate(pct_nonwhite = (pop - white)/pop, 
         area = st_area(geometry)) # area will have units

# transform to oregon planar projection
pdx_tracts_p <- st_transform(pdx_tracts, 
                             crs = 2838) %>%
  mutate(area = st_area(geometry))

# this function returns different kinds of plots depending on the input type
plot(pdx_tracts_p)

# sort tracts by area descneding
pdx_tracts_p %>%
  arrange(desc(area))

# clip two largest tracts out
pdx_tracts_p <- pdx_tracts_p %>%
  filter(area < units::as_units(1e8, "m^2"))
plot(pdx_tracts_p)

p <- pdx_tracts_p %>%
  ggplot()
p + geom_sf(aes(fill = medinc)) +
  # flip the color ramp dark = high
  # scale_fill_gradient(low = "#56B1F7", high = "#132B43") 
  # uses plasma color ramp
  scale_fill_viridis_c(option = "plasma", 
                       direction = -1
                       #, trans = "sqrt"
                       ) +
  coord_sf(datum = st_crs(2838), 
           # xlim = c(2.325e6, 2.35e6)
           ) + # crs = coordinate system, mercator, etc
              # xlim constrains x-axis
  # blank() + # clears x axis, background
  ggsn::north(pdx_tracts_p, symbol = 3, scale = 0.1) + 
  ggsn::scalebar(pdx_tracts_p, dist = 4,
                 dist_unit = "km",
                 transform = FALSE,
                 st.size = 3)
  
```

```{r next plot}
p + geom_sf(aes(fill = pct_nonwhite))

```

```{r load biketown data}
source("C:/Users/dpwpd/Documents/DataScienceClass/ds19-class/code/fetch_biketown.r")
outdir = "C:/Users/dpwpd/Documents/DataScienceClass/ds19-class/data/biketown/"
biketown <- get_data(start = "01/2018", 
                    end = "12/2018", 
                    outdir = "C:/Users/dpwpd/Documents/DataScienceClass/ds19-class/data/biketown/")

# stitch files together
biketown_2018 <- paste0("C:/Users/dpwpd/Documents/DataScienceClass/ds19-class/data/biketown/", list.files("C:/Users/dpwpd/Documents/DataScienceClass/ds19-class/data/biketown/")) %>%
  lapply(read.csv, stringsAsFactors = F) %>%
  bind_rows()

head(biketown_2018)

# for e.g. shapefiles, just use sf::read_sf()
# incl Post GIS files
biketown <- biketown_2018
hubs_start_sf <- biketown_2018 %>%
  group_by(StartHub) %>%
  summarize(lat = mean(StartLatitude),
            lng = mean(StartLongitude),
            starts = n()) %>%
  filter(!is.na(lat) & !is.na(lng)) %>%
  st_as_sf(coords = c("lng", "lat"),
           crs = 4326, # sets Coordinate System to WGS84
           agr = "constant" # don't know what this does
           )
mapview(hubs_start_sf, zcol = "starts")

hubs_end <- biketown_2018 %>%
  group_by(EndHub) %>%
   summarize(lat = mean(EndLatitude),
            lng = mean(EndLongitude),
            ends = n())

hubs_end_sf <- hubs_end %>%
  filter(!is.na(lat) & !is.na(lng)) %>%
  st_as_sf(coords = c("lng", "lat"),
          crs = 4326, # sets Coordinate System to WGS84
          agr = "constant") # don't know what this does
  
mapview(hubs_end_sf, zcol = "ends")

hubs_ratio_sf <- inner_join(hubs_start_sf,
                            hubs_end,
                            by = c("StartHub" = "EndHub")) %>%
  mutate(starts_to_ends = starts/ ends,
         ends_to_starts = ends / starts)

m1 <- mapview (hubs_ratio_sf, 
               zcol = "starts_to_ends", # sets colors
               cex = "starts_to_ends")  # sets size of dot relative to value
               # kills legend
m2 <- mapview (hubs_ratio_sf, 
               zcol = "ends_to_starts", 
               cex = "ends_to_starts")
sync(m1, m2)

```

```{r spatial joins}
hubs_start_sf_p <- hubs_start_sf %>%
  st_transform(2838)

hubs_start_sf_p <- hubs_start_sf_p %>%
  sf::st_join(pdx_tracts_p)

```
## Saving files
R data files: RDS - good for r-specific output
good to save raw data file

If it's something you want to share it with someone else who is not working in R then use write.csv

RDS files are much smaller (biketown_raw is 10 MB as RDS vs 72.5 MB as CSV)

```{r saving files}
saveRDS(pdx_tracts_raw, "C:/Users/dpwpd/Documents/DataScienceClass/ds19-class/data/pdx_tracts_raw.rds")

write.csv(biketown_2018, 
          "C:/Users/dpwpd/Documents/DataScienceClass/ds19-class/data/biketown_raw.csv",
          row.names = F)
saveRDS(biketown_2018, 
          "C:/Users/dpwpd/Documents/DataScienceClass/ds19-class/data/biketown_raw.rds")


```